---
title: "Bob Ross Data 8_6_19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

bob_ross <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-06/bob-ross.csv")
```

```{r}
bob_ross_gathered <- bob_ross %>%
      janitor::clean_names() %>% #removes extra spaces and special char
      gather(element, present, -episode, -title) %>% #wide to long
      filter(present == 1) %>%
      mutate(title = str_to_title(str_remove_all(title, '"')), #remove quotes
            element = str_to_title(str_replace(element, "_", " "))) %>%
      select(-present) %>% #remove present column after filtering
      extract(episode, c("season", "episode_number"), "S(.*)E(.*)", 
              convert = TRUE, remove = FALSE) %>%
      arrange(season, episode_number)
```

### Exploring the paintings

```{r}
bob_ross_gathered %>%
      count(element, sort = TRUE) %>%
      head(25) %>%
      mutate(element = fct_reorder(element, n)) %>%
      ggplot(aes(element, n)) + 
      geom_col() + 
      coord_flip()
```

What are the most "crowded" paintings, with the msot elements in them?

```{r}
bob_ross_gathered %>%
      add_count(episode) %>%
      arrange(desc(n))
```

How have Ross's paintings been changing over time?

```{r}
by_season_element <- bob_ross_gathered %>%
      filter(!element %in% c("Tree", "Trees")) %>%
      group_by(season) %>%
      mutate(number_episodes = n_distinct(episode)) %>%
      count(season, element, number_episodes, sort = TRUE) %>%
      mutate(percent_included = n / number_episodes) %>%
      group_by(element) %>%
      mutate(element_total = sum(n)) %>%
      ungroup()
      
by_season_element %>%
      filter(element_total >= 50) %>%
      ggplot(aes(season, percent_included, color = element)) + 
      geom_line() + 
      scale_y_continuous(labels = scales::percent_format()) +
      expand_limits(y = 0) + 
      facet_wrap(~ element)
```
### Clustering

What tends to appear together?

```{r}
library(widyr)

correlations <- bob_ross_gathered %>%
      add_count(element) %>% #remove elements with too low count
      filter(n >= 5) %>%
      pairwise_cor(element, episode, sort = TRUE)

correlations %>%
      filter(item1 == "River") %>%
      mutate(item2 = fct_reorder(item2, correlation)) %>%
      ggplot(aes(item2, correlation)) + 
      geom_col() + 
      coord_flip() + 
      labs(title = "What tends to appear with a river?", 
           subtitle = "Among elments that appear in at least 10 paintings")

correlations %>%
      filter(item1 == "Snow") %>%
      mutate(item2 = fct_reorder(item2, correlation)) %>%
      ggplot(aes(item2, correlation)) + 
      geom_col() + 
      coord_flip() + 
      labs(title = "What tends to appear with snow?", 
           subtitle = "Among elments that appear in at least 10 paintings")
```

```{r}
library(ggraph)
library(igraph)

set.seed(2019)

#create network plot
correlations %>%
      head(100) %>%
      graph_from_data_frame() %>%
      ggraph(layout = "nicely") + 
      geom_edge_link(aes(alpha = correlation)) + 
      geom_node_point() + 
      geom_node_text(aes(label = name), vjust = 1, just = 1) +
      theme_void()
```
### Principal Component Analysis (what trends predict the most)

What dimensions drive a ;lot of the variation among paintings?

```{r}
library(reshape2)
library(broom)
library(tidytext)

binary_matrix <- bob_ross_gathered %>%
      acast(title ~ element)
# Center the columns took every col and subract middle of it to balance them (37 min mark)
centered_matrix <- t(t(binary_matrix) - colMeans(binary_matrix))

svd_result <- svd(centered_matrix) 

element_weights <- tidy(svd_result, matrix = "v") %>%
      mutate(element = colnames(binary_matrix)[column])

element_weights %>%
      filter(PC <= 4) %>%
      group_by(PC) %>%
      top_n(16, abs(value)) %>%
      ungroup() %>%
      mutate(element = reorder_within(element, value, PC)) %>%
      ggplot(aes(element, value, fill = factor(PC))) + 
      geom_col(show.legend = FALSE) + 
      facet_wrap(~ PC, scales = "free") + 
      scale_x_reordered() + 
      coord_flip() + 
      labs(title = "First Four Principal Compnenets of Elements of Bob Ross Paintings")
```
1. Mountains/Vonifer vs Ocean/Beach and Deciduous Trees
2. Trees, espically deciduous vs Ocean
3. Sprint/Summer vs Winter
4. Lake vs River

```{r}
painting_weights <- broom::tidy(svd_result, matrix = "u") %>%
      mutate(painting = rownames(binary_matrix)[row]) #switch to row
```

```{r}
painting_weights %>%
      filter(PC == 1) %>%
      arrange(value)

bob_ross_gathered %>%
      filter(title == "Frozen Solitude")

painting_weights %>%
      filter(PC <= 4) %>%
      group_by(PC) %>%
      top_n(16, abs(value)) %>%
      ungroup() %>%
      mutate(painting = reorder_within(painting, value, PC)) %>%
      ggplot(aes(painting, value, fill = factor(PC))) + 
      geom_col(show.legend = FALSE) + 
      facet_wrap(~ PC, scales = "free") + 
      scale_x_reordered() + 
      coord_flip() + 
      labs(title = "First Four Principal Compnenets of Bob Ross Paintings")
```

```{r}
broom::tidy(svd_result, matrix = "d") %>%
      ggplot(aes(PC, percent)) + 
      geom_point() #lots of structure in data
```

